Refining from the BON High-level Specification Language to the VHDL-AMS Hardware Description Language

### 2009-08-21 papers/BON_VHDL/paper.tex
more text

%======================================================================
\section{Introduction}
The \STs, represents heterogeneous collection of digital, analog 
and mixed signal hardware components. Embedded system complexity 
requires the need to strive for formal descriptions each component 
separatelly in order to produce specification of the \STs simulator. 
The development of complex system requires sophisticated tool and methodology 
in order to be able to complete a correct design in time.  
The authors will illustrate in this paper how BON can represent a high 
level analogue mapping and will further illustrate a procedure for 
transformation from VHDL-AMS to BON. As an extremly simple example, 
we will give practical aspect of refinement from VHDL-AMS to BON,
applied on A/D convertor, in order to produce formal specification 
and software simulation. A complete list of the mapping between BON 
and VHDL-AMS can be found in the appendix.
%=====================================================================

@@

\begin{itemize}
\item 
\emph{Terminals} represents phisical connection points or circuit nodes 
of a network connectivity. Terminals are declared to be of various 
\emph{natures}, which represent different energy domains of a system.
\item
The \emph{simultaneous statements} creates one or more characteristic 
equations that are solved by the dedicated simulation kernel a.k.a.
"analog solver", to model continuously valued behaviour.  
\item
The \emph{quantities} denote a waveform or a time series of values.
The behaviour of the \emph{quantities} is given by the \emph{simultaneous
 statements}. They take values as a result of solving the set of simultaneous 
ordinary differential and algebraic equations (DAEs). To simplify, 
quantities can be understood as analog unknowns of (DAEs) that specify 
analog behavior. 
\item
The \emph{break statement} is an construct which allows to 
initialize the DAE system at the beginning of a transient 
simulation and to mark discontinuities during such a transient 
simulation.

\end{itemize}

\section{Choice of BON as the modeling tool}
%Brief description and motivation.
The Business Object Notation (BON) is an analysis and design notation 
for object-oriented systems  which emphasizes \emph{seamlessness}, 
\emph{reversibility} and software \emph{contracting}. Let us define 
these principles precisely and point out why the principles are 
important in VHDL-AMS simulation.
%\todo{dragan 25/07/2009 - Figure out relation: Seamlessness and VHDL-AMS}
\begin{itemize}
\item \emph{Seamlessness} is the use of a continuous process 
throughout the software lifecycle.
%\todo{dragan 25/07/2009 - Figure out relation: Reversibility and VHDL-AMS. Why is Reversibility important in VHDL design.}
\item \emph{Reversibility} is the support for both forward and 
backward development process: from analysis to design 
and implementation, and back.
In VHDL-AMS, modelling represents an continious top-down-bottom-up
process. Top-down approach serves for design while bottom-up
serves for verification. The possibility of seamlessly translating 
changes made during a certain development phase back into earlier phases,
facilitates a stable and predictable environment and ensures the various 
models consistent with the code.
\item \emph{Contracting} is the precise definition, for each software element, 
of the associated semantic properties. BON use a full assertion mechanism, 
allowing analysts to specify not only the structure of a system but also its
semantics ( constraints and invariants of the expected results ). 
In VHDL-AMS \emph{simultaneous statement} represents analogue constraint of 
continuous-time waveform. In other words, simulator evaluates quantities such 
that the constraints specified by the simultaneous statements are satisfied. 
We admit that the constraint of waveform is nothing else than \emph{invariant}.
Another example of \emph{invariant} is conservation of energy law in which the summation
of \arcoss quantities (\texttt{$\sum_{loop}$ across quatities =0}) 
around any closed path (loop) equals zero and the summation of
through quantities (\texttt{$\sum_{node}$ through quatities =0}) at any node (point) equals zero.
Finally, any predefined (implicite) attribute in VHDL has defined \emph{pre}/\emph{(post)condition} trought
input/output parameters and theirs constraints.

\end{itemize}

>>>>

%======================================================================
\section{Introduction}
The \STs, represents heterogeneous collection of digital, analog 
and mixed signal hardware components. Embedded system complexity 
requires the need to strive for formal descriptions each component 
separatelly in order to produce specification of the \STs simulator. 
The development of complex system requires sophisticated tool and methodology 
in order to be able to complete a correct design in time.  
The authors will illustrate in this paper how BON can represent a high 
level analogue mapping and will further illustrate a procedure for 
transformation from VHDL-AMS to BON. As an extremly simple example, 
we will give practical aspect of refinement from VHDL-AMS to BON,
applied on A/D convertor, in order to produce formal specification 
and software simulation. A comprehensive list of the mapping between BON 
and VHDL-AMS can be found in the appendix.
%=====================================================================

@@

\begin{itemize}
\item 
\emph{Terminals} represents phisical connection points or circuit nodes 
of a network connectivity. They are declared to be of various 
\emph{natures}, which represent different energy domains of a system.
Terminals hold no values; their only role is to facilitate the formation
of conservative equation sets that, in turn, constrain the associated 
\emph{branch quatities}. 
\item
The \emph{simultaneous statements} creates one or more characteristic 
equations that are solved by the dedicated simulation kernel a.k.a.
"analog solver", to model continuously valued behaviour.  
\item
The \emph{quantities} denote a waveform or a time series of values.
The behaviour of the \emph{quantities} is given by the \emph{simultaneous
 statements}. They take values as a result of solving the set of simultaneous 
ordinary differential and algebraic equations (DAEs). To simplify, 
quantities can be understood as analog unknowns of (DAEs) that specify 
analog behavior. There are three kinds of quantity in VHDL-AMS: free, 
branch and source quantities. A \emph{free quantity} can be used in 
signal-flow modeling. A \emph{branch quantity} is used to model conservative energy
systems. A \emph{source quantity} is used for frequency and noise modeling.
\item
The \emph{break statement} is an construct which allows to 
initialize the DAE system at the beginning of a transient 
simulation and to mark discontinuities during such a transient 
simulation. A break statement allows restart of the analog kernel,
than overrides quantity values and executes concurrently with the analogue model. 
For any concurrent break statement there is an associated process.
\end{itemize}

\section{Choice of BON as the modeling tool}
%Brief description and motivation.
The Business Object Notation (BON) is an analysis and design notation 
for object-oriented systems  which emphasizes \emph{seamlessness}, 
\emph{reversibility} and software \emph{contracting}. Let us define 
these principles precisely and point out why the principles are 
important in VHDL-AMS simulation.

\subsection{Seamlessness and Reversibility}
\emph{Seamlessness} is the use of a continuous process 
throughout the software lifecycle.
\emph{Reversibility} is the support for both forward and 
backward development process: from analysis to design 
and implementation, and back.
In VHDL-AMS, modelling represents an continious top-down-bottom-up
process. Top-down approach serves for design while bottom-up
serves for verification. The possibility of seamlessly translating 
changes made during a certain development phase back into earlier phases,
facilitates a stable and predictable environment and ensures the various 
models consistent with the code. Keeping in sync various models, documetation
and specification from forward progress to backward and redesign, preclude 
that design and implementation diverge to an inconsistent state.   
\subsection{Contracting}
\emph{Contracting} is the precise definition, for each software element, 
of the associated semantic properties. BON use a full assertion mechanism, 
allowing analysts to specify not only the structure of a system but also its
semantics ( constraints and invariants of the expected results ). 
In VHDL-AMS \emph{simultaneous statement} represents analogue constraint of 
continuous-time waveform. In other words, simulator evaluates quantities such 
that the constraints specified by the simultaneous statements are satisfied. 
We admit that the constraint of waveform is nothing else than \emph{invariant}.
Another example of \emph{invariant} is conservation of energy law in which 
the summation of \arcoss quantities (\texttt{$\sum_{loop}$ across quatities =0}) 
around any closed path (loop) equals zero and the summation of through 
quantities (\texttt{$\sum_{node}$ through quatities =0}) at any node 
(point) equals zero. Finally, any predefined (implicite) attribute in VHDL 
has defined \emph{pre}/\emph{post condition} trought input/output 
parameters and theirs constraints.
